### 2014-09-06 eval.tex
relax I typos fixed
citations fixed

\section{Evaluation}~\label{sec:eval}
Our evaluation focuses on the effectiveness  and scalability.
To measure the effectiveness, we compare with the predictive analysis, 
{\sf RV}~\cite{}. We chose RV for two main reasons: (1) {\sf RV} is the 
only open source predictive analysis, (2) {\sf RV} represents the 
state-of-the-art, which is theoretically proven to have higher detection 
capability than other approaches.

@@

To improve the scalability in constraint solving, we again apply 
a splitting optimization. The full trace is split into $N$ equal-length 
subtraces or ``time windows'' (in our experiments, $N=1,000$). The values 
of local and shared variables are then passed at the end of a 
given window as the inputs to the next window. For lack of space, we 
omit the details, and instead refer the reader to 
Huang et al. \cite{HuangMR14}, who apply a similar optimization.

{\bf Evaluation Method} We conducted our experiments on a set of 16 
applications, which were also used to evaluate {\sf RV}. Specifically, 
the set includes large applications such as {\sf Jigsaw}, 
{\sf Xalan} and {\sf Lusearch}. To handle large applications that make 
use of reflection, we applied the {\sf Tamiflex} tool~\cite{}, which 
replaces reflective calls with the concrete method invocations recorded 
in the observed run.
% for the purpose of effective static analysis.  
We omit the benchmark {\sf eclipse} because the current version 
of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, 
which throws an exception when the main service is started. 
By applying \tool to the abnormal execution, we identify only 3 races, 
similarly to the report of RV~\cite{}.
In addition, the {\sf montecarlo} benchmark requires an external input 
that we downloaded from the internet and simplified. For the large 
applications, we utilized the most lightweight available configuration.

@@

First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}. There 
are several reasons for that: (1) {\sf Recipe} can reason about the 
accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can 
only reason about the accesses in the executed paths. (2) {\sf Recipe} 
or {\sf Recipe-s} relax the scheduling even if it 
%breaks the inter-thread read/write dependence structure 
changes variable values in the observed run. {\sf Recipe} allows the 
majority of the shared reads, i.e., the reads of primitive values, 
to freely read a different value from a different write 
as long as the value leads to the same branch decisions. {\sf RV}, however, 
requires them to read the same values as in the observed run. (3) An 
critical optimization proposed by {\sf RV} is to preserve dependences 
only for the reads before the preceding branches of the racy events, 
rather than all the reads. However, this optimization is underplayed 
by the fact that {\sf RV} introduces huge amount of artificial 
branches, i.e., one before each shared field access, which ensures 
the use of the same base objects.
We get rid of such artificial branches and instead, rely on the small 
amount of base read events to ensure the use of the same base 
objects (Section~\ref{}). Our strategy reduces the number of branches 
greatly and amplifies the effectiveness of the optimization. We also 
carried out case studies (Section~\ref{}) to better illustrate the scenarios.

@@

\subsection{Case studies}
We manually inspected the reported races over small applications to gain better understanding about our approach.

{\bf Relaxing Values\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
The code is from the benchmark {\sf bbuffer}, where the line numbers are
marked. Huang et al.~\cite{} detect the race between lines 291 and 400, 
but fail to detect the race between lines 294 and 400. The reason is 
that in the observed run, the execution follows the order 400, 291 and 294. 
For the event at line 294, its preceding branch at line 291 reads 
from line 400. Therefore, Huang et al.~\cite{}
require the predicted run to preserve the dependence between line 291 and 
line 400, so that line 291 reads exactly the same value and the branching 
expression evaluates to the same truth value. The dependence enforces 
the order constraint, $400 \rightarrow 291$, which further enforces 
the order $400\rightarrow 291 \rightarrow 294$.    Our approach does not 
require the existence of such dependence. Specifically, we allow line 
291 to happen before line 400 in the predicted run as long as the value 
read by it leads to the same branch decision, which is true in this case.  
As a result, there is no order constraint between line 294 and 
line 400, and the two form a race. We were able to replay this race  
easily using the eclipse IDE breakpoints.

@@

Initially there is one sorting thread. After it starts Thread 1 and 
Thread 2, there are three threads alive and only two more 
entries are available. Suppose that in the observed run Thread 2 
consumes both thread entries and starts the child Thread 3 
and Thread 4 (not shown), updating {\tt alive} to 5. Then Thread 1 
cannot execute the branch at lines 4-5. Huang et al.~\cite{} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2. As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5. Our analysis does not suffer from this limitation. Instead, it allows the predicted run to reason about the un-explored code. Specifically, it does not need to preserve the dependence from line 11 to line 1, and it allows line 1 (Thread 1) to read from line 9 (Thread 2). As a result, the branch condition guarding the unexplored branch is evaluated to be true, enabling the unexplored path in the constraint solver. Finally, the solver identifies the race between line 5 (Thread 1) and line 1 (Thread 3). Note that the two lines are synchronized on different locks~\footnote{We abbreviate the {\tt synchronized} keyword as {\tt sync}.
}.

>>>>

\section{Evaluation}~\label{sec:eval}
Our evaluation focuses on the effectiveness  and scalability.
To measure the effectiveness, we compare with the predictive analysis, 
{\sf RV}~\cite{pldi14}. We chose RV for two main reasons: (1) {\sf RV} is the 
only open source predictive analysis, (2) {\sf RV} represents the 
state-of-the-art, which is theoretically proven to have higher detection 
capability than other approaches.

@@

To improve the scalability in constraint solving, we again apply 
a splitting optimization. The full trace is split into $N$ equal-length 
subtraces or ``time windows'' (in our experiments, $N=1,000$). The values 
of local and shared variables are then passed at the end of a 
given window as the inputs to the next window. For lack of space, we 
omit the details, and instead refer the reader to 
Huang et al. \cite{pldi14}, who apply a similar optimization.

{\bf Evaluation Method} We conducted our experiments on a set of 16 
applications, which were also used to evaluate {\sf RV}. Specifically, 
the set includes large applications such as {\sf Jigsaw}, 
{\sf Xalan} and {\sf Lusearch}. To handle large applications that make 
use of reflection, we applied the {\sf Tamiflex} tool~\cite{tamiflex}, which 
replaces reflective calls with the concrete method invocations recorded 
in the observed run.
% for the purpose of effective static analysis.  
We omit the benchmark {\sf eclipse} because the current version 
of {\sf Tamiflex}  leads to abnormal execution after the instrumentation, 
which throws an exception when the main service is started. 
By applying \tool\  to the abnormal execution, we identify only 3 races, 
similarly to the report of RV~\cite{pldi14}.
In addition, the {\sf montecarlo} benchmark requires an external input 
that we downloaded from the internet and simplified. For the large 
applications, we utilized the most lightweight available configuration.

@@

First of all, {\sf Recipe} finds $>$150 more races than {\sf RV}. There 
are several reasons for that: (1) {\sf Recipe} can reason about the 
accesses in the un-executed paths, while {\sf RV} and {\sf Recipe-s} can 
only reason about the accesses in the executed paths. (2) {\sf Recipe} 
or {\sf Recipe-s} relax the scheduling even if it 
%breaks the inter-thread read/write dependence structure 
changes variable values in the observed run. {\sf Recipe} allows the 
majority of the shared reads, i.e., the reads of primitive values, 
to freely read a different value from a different write 
as long as the value leads to the same branch decisions. {\sf RV}, however, 
requires them to read the same values as in the observed run. (3) An 
critical optimization proposed by {\sf RV} is to preserve dependences 
only for the reads before the preceding branches of the racy events, 
rather than all the reads. However, this optimization is underplayed 
by the fact that {\sf RV} introduces huge amount of artificial 
branches, i.e., one before each shared field access, which ensures 
the use of the same base objects.
We get rid of such artificial branches and instead, rely on the small 
amount of base read events to ensure the use of the same base 
objects (Section~\ref{sec:relax1} and Section~\ref{sec:relax2}). Our strategy reduces the number of branches 
greatly and amplifies the effectiveness of the optimization. We also 
carried out case studies (Section~\ref{sec:cases}) to better illustrate the scenarios.

@@

\subsection{Case studies}~\label{sec:cases}
We manually inspected the reported races over small applications to gain better understanding about our approach.

{\bf Relaxing Values\ } Figure~\ref{fig:relax1} demonstrates the relaxation of inter-thread dependence enabled by our approach. 
The code is from the benchmark {\sf bbuffer}, where the line numbers are
marked. Huang et al.~\cite{pldi14} detect the race between lines 291 and 400, 
but fail to detect the race between lines 294 and 400. The reason is 
that in the observed run, the execution follows the order 400, 291 and 294. 
For the event at line 294, its preceding branch at line 291 reads 
from line 400. Therefore, Huang et al.~\cite{pldi14}
require the predicted run to preserve the dependence between line 291 and 
line 400, so that line 291 reads exactly the same value and the branching 
expression evaluates to the same truth value. The dependence enforces 
the order constraint, $400 \rightarrow 291$, which further enforces 
the order $400\rightarrow 291 \rightarrow 294$.    Our approach does not 
require the existence of such dependence. Specifically, we allow line 
291 to happen before line 400 in the predicted run as long as the value 
read by it leads to the same branch decision, which is true in this case.  
As a result, there is no order constraint between line 294 and 
line 400, and the two form a race. We were able to replay this race  
easily using the eclipse IDE breakpoints.

@@

Initially there is one sorting thread. After it starts Thread 1 and 
Thread 2, there are three threads alive and only two more 
entries are available. Suppose that in the observed run Thread 2 
consumes both thread entries and starts the child Thread 3 
and Thread 4 (not shown), updating {\tt alive} to 5. Then Thread 1 
cannot execute the branch at lines 4-5. Huang et al.~\cite{pldi14} require the predicted run to preserve the dependence denoted by the dotted line since this dependence affects the branch condition at line 2. As a result, the predicted run follows the same branch decision and cannot reason about the code at lines 4-5. Our analysis does not suffer from this limitation. Instead, it allows the predicted run to reason about the un-explored code. Specifically, it does not need to preserve the dependence from line 11 to line 1, and it allows line 1 (Thread 1) to read from line 9 (Thread 2). As a result, the branch condition guarding the unexplored branch is evaluated to be true, enabling the unexplored path in the constraint solver. Finally, the solver identifies the race between line 5 (Thread 1) and line 1 (Thread 3). Note that the two lines are synchronized on different locks~\footnote{We abbreviate the {\tt synchronized} keyword as {\tt sync}.
}.

### 2014-09-06 overview.tex
citations fixed

\smallskip
\noindent
{\em I. Value Relaxation\ } Our first relaxation is to allow variables
to have different values as long as the same thread local path is preserved.
It does not require data dependences or same variable values as in~\cite{yannis, pldi14, Said:2011}.
This allows us to catch the racy pair $(e_2, e_5)$.
%A potential race is between 
%the pair, $(e_2, e_5)$, as they access the shared variable $x$ from 
%different threads and one is a write. 
Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
 O, \theta>$, our analysis computes a new trace,  
$\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
 $\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still 
executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules 
the events from different threads so that the potential race pair runs 
concurrently, i.e., $e_2$ and $e_5$ run concurrently. When the schedule 
changes, the variables may get different values. Therefore, we need to 
compute both the new schedule $O'$ and the new value mapping $\theta'$, in 
order to witness a race. To guarantee that $\tau'$ is a feasible trace of 
the program $P$, the computation needs to respect a set of constraints, 
which include the value constraints, control flow constraints and 
synchronization constraints~\footnote{We omit the synchronization constraints 
in this paper, which are already well explained in existing 
techniques~\cite{yannis, pecan}.}. We explain the constraints 
briefly in the following. Details are in Section~\ref{sec:relax1} 
and the soundness guarantee is explained in Section~\ref{sec:guarantee}.

>>>>

\smallskip
\noindent
{\em I. Value Relaxation\ } Our first relaxation is to allow variables
to have different values as long as the same thread local path is preserved.
It does not require data dependences or same variable values as in~\cite{yannis, pldi14, Said:2011}.
This allows us to catch the racy pair $(e_2, e_5)$.
%A potential race is between 
%the pair, $(e_2, e_5)$, as they access the shared variable $x$ from 
%different threads and one is a write. 
Given a trace, $\tau=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \},
 O, \theta>$, our analysis computes a new trace,  
$\tau'=<\Gamma , \{\tau_{t_1}, \tau_{t_2}, \dots \tau_{t_n} \}, O', \theta'>$.
 $\tau'$ preserves the same event sequence for each thread, i.e., $T_1$ still 
executes $e_1 e_2 e_3$ and $T_2$ still executes $e_4 e_5$, but it reschedules 
the events from different threads so that the potential race pair runs 
concurrently, i.e., $e_2$ and $e_5$ run concurrently. When the schedule 
changes, the variables may get different values. Therefore, we need to 
compute both the new schedule $O'$ and the new value mapping $\theta'$, in 
order to witness a race. To guarantee that $\tau'$ is a feasible trace of 
the program $P$, the computation needs to respect a set of constraints, 
which include the value constraints, control flow constraints and 
synchronization constraints~\footnote{We omit the synchronization constraints 
in this paper, which are already well explained in existing 
techniques~\cite{yannis, pldi14}.}. We explain the constraints 
briefly in the following. Details are in Section~\ref{sec:relax1} 
and the soundness guarantee is explained in Section~\ref{sec:guarantee}.

### 2014-09-06 related.tex
citations fixed

Said et al. \cite{Said:2011} and Huang et al. \cite{HuangMR14} proposed
PTAs that are also sound. Said et al. perform symbolic analysis of the 
input trace, and then utilize an SMT solver to search for schedules that 
establish the presence of data races. Soundness is guaranteed by their 
ability to precisely encode the semantics of sequential consistency. 
Inspired by Said et al., Huang et al. present an improvement, whereby 
control-flow information is encoded into the constraint system so as 
to relax flow dependencies as long as control dependencies are 
preserved. ExceptionNULL \cite{Farzan:2012} is another example of 
sound PTA. Its goal is to detect null dereferences rather than data races.

@@

\subsection{Maximality Claim}
Serbanuta et al. \cite{maximal} and Huang et al.~\cite{HuangMR14} both 
claim maximality of the detection capability, but under different assumptions. 
Assuming branch events are not modeled in the trace, \cite{maximal} guarantees maximality. Given few 
information in the trace, it conservatively requires the shared read to 
obtain the same value in the predicted run to ensure soundness. However, if 
branch events are modeled, the technique does not guarantee maximality.  
Assuming the branch events are modeled but local assignments are not recorded, 
\cite{HuangMR14} guarantees maximality. 
Compared to \cite{maximal}, it allows some events to read different 
values if they do not affect the following branch decisions. However, 
it requires the events that affect the branch decisions to read exactly 
the same values. In practice, these assumptions are too restrictive  
because local assignments can be easily collected in practice.
%the testers often have full access to the source/bytecode and 
%can collect local assignments freely.

>>>>

Said et al. \cite{Said:2011} and Huang et al. \cite{pldi14} proposed
PTAs that are also sound. Said et al. perform symbolic analysis of the 
input trace, and then utilize an SMT solver to search for schedules that 
establish the presence of data races. Soundness is guaranteed by their 
ability to precisely encode the semantics of sequential consistency. 
Inspired by Said et al., Huang et al. present an improvement, whereby 
control-flow information is encoded into the constraint system so as 
to relax flow dependencies as long as control dependencies are 
preserved. ExceptionNULL \cite{Farzan:2012} is another example of 
sound PTA. Its goal is to detect null dereferences rather than data races.

@@

\subsection{Maximality Claim}
Serbanuta et al. \cite{maximal} and Huang et al.~\cite{pldi14} both 
claim maximality of the detection capability, but under different assumptions. 
Assuming branch events are not modeled in the trace, \cite{maximal} guarantees maximality. Given few 
information in the trace, it conservatively requires the shared read to 
obtain the same value in the predicted run to ensure soundness. However, if 
branch events are modeled, the technique does not guarantee maximality.  
Assuming the branch events are modeled but local assignments are not recorded, 
\cite{pldi14} guarantees maximality. 
Compared to \cite{maximal}, it allows some events to read different 
values if they do not affect the following branch decisions. However, 
it requires the events that affect the branch decisions to read exactly 
the same values. In practice, these assumptions are too restrictive  
because local assignments can be easily collected in practice.
%the testers often have full access to the source/bytecode and 
%can collect local assignments freely.
